---
title: "Hands-on Exercise 1"
author: "Magdalene Chan"
date: "`r Sys.Date()`"
---

# Getting Started

## Installation and launch of R packages

The code chunk below uses p_load() of pacman package to check if tidyverse packages have been installed on the computer. If they are, the packages will be launched.

```{r}
pacman::p_load(tidyverse)
```

## Importing the data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

# Plotting a simple bar chart

```{r}
ggplot(data = exam_data,
       aes(x = RACE)) + 
  geom_bar()
```

# Working with facets

## Ploting trellis boxplots

```{r}
ggplot(data = exam_data,
       aes(x = CLASS, y = MATHS)) + 
  geom_boxplot(outlier.colour="black",
               outlier.shape=16,
               outlier.size=1, 
               notch=FALSE) + 
  facet_grid(cols = vars(GENDER))
```

```{r}
ggplot(data = exam_data,
       aes(x = CLASS, y = MATHS)) + 
  geom_boxplot(outlier.colour="black",
               outlier.shape=16,
               outlier.size=1, 
               notch=FALSE) + 
  facet_grid(rows = vars(GENDER))
```

```{r}
ggplot(data = exam_data,
       aes(x = GENDER, y = MATHS)) + 
  geom_boxplot(outlier.colour="black",
               outlier.shape=16,
               outlier.size=1, 
               notch=FALSE) + 
  facet_grid(cols = vars(CLASS),
             rows = vars(GENDER))
```

# Working with themes

## Ploting horizontal bar chart

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "light blue",
                                        colour = "light blue"),
        panel.grid.major = element_line(linewidth = 0.5,
                                        linetype = 'solid',
                                        colour = "white"),
        panel.grid.minor = element_line(linewidth = 0.25,
                                        linetype = 'solid',
                                        colour = "white"))
```

# Designing Data-drive Graphics for Analysis

## Makeover design

```{r}
pacman::p_load(forcats)

exam_data %>%
  mutate(RACE = fct_infreq(RACE)) %>%
  ggplot(aes(x=RACE)) +
  geom_bar() + 
  ylim(0,225) + 
  geom_text(
    stat = "count",
    aes(label=paste0(after_stat(count),", ",
                     round(after_stat(count)/sum(after_stat(count))*100,
                           1),"%")),
                      vjust=-1,
                      size=3) +
  xlab("Race") +
  ylab("No. of\nPupils") +
  theme(axis.title.y = element_text(angle=0))

```

```{r}
ggplot(data=exam_data, 
       aes(x=MATHS)) +
  geom_histogram(bins = 20, 
                 color = "black", 
                 fill = "light blue") + 
  geom_vline(aes(xintercept = mean(MATHS)),
             linetype = "dashed",
             color = 'red') +
  geom_vline(aes(xintercept = median(MATHS)),
             linetype = "dashed",
             color = 'black')

```

```{r}
eng_bg <- exam_data[,-3]

ggplot(data=exam_data, 
       aes(x=ENGLISH, fill=GENDER)) +
  geom_histogram(data = eng_bg, fill = "grey") +
  geom_histogram(color = "black") +
  facet_wrap(~GENDER) +
  ylim(0,30)

```

```{r}
ggplot(data=exam_data, 
       aes(x=MATHS, y=ENGLISH)) +
  geom_point(size=1) +
  xlim(0,105) +
  ylim(0,105) +
  geom_vline(aes(xintercept = 50),
             linetype = "dashed",
             color = 'grey') +
  geom_hline(aes(yintercept = 50),
             linetype = "dashed",
             color = 'grey')

```
